### 2020-09-29 paper/natastron/paper.tex

\documentclass[12pt]{article}
\setlength{\textwidth}{167mm}
\setlength{\headsep}{0mm}
\setlength{\headheight}{0mm}
\setlength{\textheight}{230mm}
\setlength{\oddsidemargin}{1.6mm}
\setlength{\evensidemargin}{1.6mm}
\setlength{\topmargin}{4.6mm}

\usepackage{chngcntr}
\usepackage{xspace}
\usepackage{graphicx}
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage[ ]{hyperref}
\graphicspath{{Images/}}
\usepackage[symbol]{footmisc}

% Astronomical abbreviations (thanks to Dan Huber)
\newcommand{\numax}{\mbox{$\nu_{\rm max}$}\xspace}
\newcommand{\dnu}{\mbox{$\Delta \nu$}\xspace}
\newcommand{\muHz}{\mbox{$\mu$Hz}\xspace}
\newcommand{\teff}{\mbox{$T_{\rm eff}$}\xspace}
\newcommand{\logg}{\mbox{$\log(g)$}\xspace}
\newcommand{\feh}{\mbox{$\rm{[Fe/H]}$}\xspace}
\newcommand{\msun}{\mbox{$\mathrm{M}_{\odot}$}\xspace}
\newcommand{\lsun}{\mbox{$\mathrm{L}_{\odot}$}\xspace}
\newcommand{\mearth}{\mbox{$\mathrm{M}_{\oplus}$}\xspace}
\newcommand{\rsun}{\mbox{$\mathrm{R}_{\odot}$}\xspace}
\newcommand{\kepler}{\emph{Kepler}\xspace}
\newcommand{\tess}{\emph{TESS}\xspace}
\newcommand{\ktwo}{K2\xspace}
\newcommand{\gaia}{\emph{Gaia}\xspace}

\begin{document}
\noindent\textbf{\LARGE{New asteroseismic rotation rates of \emph{Kepler} dwarfs show strong agreement with weakened magnetic braking on the late-age main sequence}}\\

\noindent Oliver J. Hall$^{1,2,3,}$\footnote[1]{Email: \texttt{oliver.hall@esa.int} | Twitter: \texttt{@asteronomer} | GitHub: \texttt{@ojhall94}}
	Guy R. Davies$^{2,3}$, 
	Jennifer van Saders$^{4,5,6}$,
	Martin B. Nielsen$^{2,3}$
	Mikkel N. Lund$^{3}$, 
	William J. Chaplin$^{2,3}$, 
	Rafael A. Garc\'ia$^{7, 8}$, 
	Louis Amard$^{9}$,
	Angela A. Breimann$^{9}$, 
	Saniya Khan$^{2,3}$, 
	Victor See$^{9}$, 
	Jamie Tayar$^{4, 10}$
	\\

	% List of institutions
	\noindent $^{1}$ European Space Agency (ESA), European Space Research and Technology Centre (ESTEC), Keplerlaan 1, 2201 AZ Noordwijk, The Netherlands

	\noindent 	$^{2}$ School of Physics and Astronomy, University of Birmingham, Edgbaston, Birmingham, B15 2TT, UK

	\noindent 	$^{3}$ Stellar Astrophysics Centre, Department of Physics and Astronomy, Aarhus University, Ny Munkegade 120, 8000 Aarhus C, Denmark

	\noindent 	$^{4}$ Institute for Astronomy, University of Hawai'i, Honolulu, HI 96822

	\noindent 	$^{5}$ Observatories of the Carnegie Institution for Science, Pasadena, CA 91101

	\noindent 	$^{6}$ Department of Astrophysical Sciences, Princeton University, Princeton, NJ 08544

	\noindent 	$^{7}$ IRFU, CEA, Universit\'e Paris-Saclay, F-91191 Gif-sur-Yvette, France

	\noindent 	$^{8}$ AIM, CEA, CNRS, Universit\'e Paris-Saclay, Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, F-91191 Gif-sur-Yvette, France

	\noindent 	$^{9}$ University of Exeter Department of Physics and Astronomy, Stocker Road, Devon, Exeter, EX4 4QL, UK

	\noindent $^{10}$ Hubble Fellow

\vspace{10mm}

\textbf{Studies using asteroseismic ages and rotation rates from star-spot rotation have indicated that standard age-rotation relations may break down roughly half-way through the main sequence lifetime, a phenomenon referred to as weakened magnetic braking. While rotation rates from spots can be difficult to determine for older, less active stars, rotational splitting of asteroseismic oscillation frequencies can provide rotation rates for both active and quiescent stars, and so can confirm whether this effect really takes place in older stars.\\
We obtained asteroseismic rotation rates of 91 main sequence stars showing high signal-to-noise modes of oscillation.
Using these new rotation rates, along with effective temperatures, metallicities and seismic masses and ages, we built a hierarchical Bayesian mixture model to determine whether the ensemble more closely agreed with a standard rotational evolution scenario \cite{vansaders+pinsonneault2013}, or one where weakened magnetic braking takes place \cite{vansaders+2016}. The weakened magnetic braking scenario was found to be $98.4\%$ more likely for our stellar ensemble, adding to the growing body of evidence for this stage of stellar rotational evolution. This work represents the largest catalogue of seismic rotation on the main sequence to date, opening up possibilities for more detailed ensemble analysis of rotational evolution with \textit{Kepler}.}

%%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%
 %500 words
Gyrochronology is the study of the relationship between a star's rotation period and its age. As a star grows older along the main sequence, magnetic winds will cause it to lose angular momentum, slowing it down. Because the loss rate is related to temperature, the rotation period of a young star will rapidly settle on to a plane in age-colour-rotation space \cite{barnes2007}. As a result, knowing the rotation and colour of a star provides an avenue to measure it's age, to a precision of $\sim 10\%$ for stars most similar to the Sun \cite{meibom+2015}. Age can otherwise be difficult to come by, and so gyrochronology enables more in-depth studies of stellar populations and individual stars \cite{leiner+2019,claytor+2019}.

Gyrochronology was previously calibrated on stellar clusters, which have well constrained ages, but only up to roughly $2.5\, \rm{Gyr}$ so far \cite{meibom+2015}. With the launch of the \textit{Kepler} mission \cite{borucki+2010}, ages of main sequence field stars (i.e. not in clusters) became more widely available through asteroseismology, the study of stellar variability \cite{silvaaguirre+2015}. When looking at these stars, disagreements were found between the observations and gyrochronology beyond the middle of stars' main sequence lifetimes, which could not be reconciled with existing theories \cite{angus+2015, nielsen+2015, davies+2015}. It was proposed that at some stage in a star's evolution it undergoes \textit{weakened magnetic braking} (WMB, \cite{vansaders+2016}), where the efficiency of angular momentum loss rapidly drops. This would cause stars to keep fast rotation rates at old ages, which we would not expected from existing gyrochronology relations.

The mechanism by which weakened magnetic braking occurs has since been the subject of ongoing debate, and may be connected to changes in the magnetic field morphology \cite{reville+2015,garraffo+2016, metcalfe+2019, see+2019}. It may also be explained from an observational point of view. A large scale survey of stellar rotation rates measured using star-spots pointed out a lack of slowly rotating stars older then Sun \cite{mcquillan+2014}. As activity reduces with age, older stars with fewer star-spots are less likely to have their rotation measured \cite{matt+2015}. The point at which the detection probability drops appears to lie at a similar level of activity as when the proposed weakened magnetic braking takes place, pitting these two possibilities against one another \cite{vansaders+2019}.

Determining whether weakened magnetic braking is a true phenomenon or a bias in star-spot observations requires new data, which can be provided by asteroseismology. A star's rotation causes modes of oscillation to split into multiplets, making it possible to obtain an independent rate of rotation by measuring the splitting between oscillation frequencies \cite{ledoux1951}. Asteroseismic rotation is more difficult to obtain than spot rotation, but does not require spots to be visible, meaning that we can measure rotation for older, quiescent stars that would not have been present in existing rotation catalogues. This makes asteroseismic rotation rates the perfect avenue to explore whether weakened magnetic braking truly takes place on the main sequence. Here, we measured asteroseismic rotation rates of 91 main sequence stars across a broad range of  ages, and evaluated this new ensemble to determine whether it agrees more closely with a classical rotational evolution scenario, or one where weakened magnetic braking takes place.\\

\textbf{Asteroseismic main sequence targets:} In order to obtain robust asteroseismic rotation rates, we required detections of multiples of both dipole (denoted as $\ell = 1$) and quadrupole ($\ell = 2$) oscillations for each star, of which the latter have significantly lower signal-to noise. Radial oscillations ($\ell = 0$), which are also visible, do not split. In this work, we studied a sample of 95 of the highest signal-to-noise targets observed by \textit{Kepler}, combining the `Kages' \cite{silvaaguirre+2015,davies+2016} and LEGACY \cite{lund+2017, silvaaguirre+2017} catalogues\footnote{For these catalogues the mode extraction through frequency fitting \cite{davies+2016, lund+2017} and modelling using mode frequencies to obtain stellar parameters \cite{silvaaguirre+2015, silvaaguirre+2017} are covered in separate papers.}. Further details on how these data can be found in \textbf{\textit{Methods}}.

We separated our sample into three categories \cite{garcia+2014}: 64 main sequence stars (MS, $\rm{log(g)} > 4.0\, \rm dex$, $T_{\rm eff} < 6250\, \rm K$); 4 potential sub-giant stars (SG, $\rm{log(g)} < 4.0\, \rm dex$, $T_{\rm eff} < 6250\, \rm K$), which may have begun to evolve off the main sequence, complicating their rotational history\footnote{It should be noted that none of these stars have identified mixed dipole modes, which indicate an evolved structure (see \cite{bedding+2010}).}; and 24 `hot' stars (H, $T_{\rm eff} > 6250\, \rm K$), which lie above the Kraft break, the point at which convective envelopes are tenuous, and angular momentum loss through magnetic winds becomes inefficient \cite{kraft1967}. Our conclusions are largely insensitive to these categorical assignments, and they are intended solely to help explore results category-by-category. The sample, which can be seen in Figure \ref{fig:sample}, spans surface gravities of $3.8\, \rm{dex} < \rm{log(g)} < 4.6\, \rm{dex}$, temperatures of $5000\, \textrm{K} < T_{\rm eff} <  6700\, \rm{K}$, and ages from $1$ to $\sim 13\, \rm{Gyr}$.\\

\begin{figure}
	\centering
	\includegraphics[width=0.6\textwidth]{data.pdf}
	\caption{Our sample of 95 stars from the Kages and LEGACY catalogues coloured by asteroseismic age  \cite{silvaaguirre+2015, silvaaguirre+2017}. The dashed lines indicate our classification boundaries: main sequence (circles, $\teff < 6250\, K$, $\logg < 4\, \rm{dex}$), potential sub-giants (squares, $\teff < 6250\, K$, $\logg > 4\, \rm{dex}$), and `hot' stars (triangles, $\teff > 6250\, K$). The Sun is shown with the `$\odot$' symbol, and has an age of $4.6\, \rm{Gyr}$ \cite{bonanno+frohlich2015}. The solid lines are evolutionary tracks generated using MESA \cite{paxton+2017}, for a metallicity of $Z = 0.01493$ and helium content of $Y = 0.26588$. Left to right, they represent masses of $1.5$, $1.25$, $1$ and $0.75\, M_\odot$. Vertical errorbars are too small to be visible.}
	\label{fig:sample}
\end{figure}

\textbf{Results of asteroseismic fitting:} The stars in `Kages' and LEGACY form our stellar sample have been subject to detailed asteroseismic frequency fitting in the catalogue papers, but did not report asteroseismic rotation. Here, we repeated the mode frequency fitting with a Bayesian model that accounted for rotation in more detail, and improved the treatment of existing asteroseismic relations in frequency fitting techniques (see \textbf{\textit{Methods}} and \textit{\textbf{Supplementary Information}}). We fit our new model to \kepler power-spectra of 95 stars, which was successful in 94 cases. We compared our results with both published and unpublished asteroseismic rotation rates, and found that the three stars with the lowest angles of inclination ($< 10^\circ$) were inconsistent with multiple other studies. To some degree this was expected, as  for stars viewed near pole-on the split modes of oscillation have extremely low power, making them hard to detect reliably. We flagged our rotation rates for these stars, and did not use them in the next steps in our analysis, leaving a sample of 91 stars for which we considered our asteroseismic rotation measurements to be robust (for a full justification, see \textit{\textbf{Methods}}).\\

\textbf{Seismic versus surface rotation rates:} Different methods of rotation measurement probe different depths of stars. Asteroseismology of main sequence stars probes internal rotation in the near surface layers, where the observed modes of oscillation are most sensitive \cite{lund+2014}. Measurements of star-spot modulation instead probe the rotation rates of star-spots on the surface. Previous studies have shown that seismic rotation rates show no statistically significant deviation from spot rotation rates \cite{nielsen+2015, benomar+2015}. We confirm that this holds true for our larger sample, in order to ensure that our seismic rotation periods can be reliably used to draw conclusions on the evolution of surface rotation.

Figure \ref{fig:protlit} shows a comparison between 48 stars from our sample for which spot  rotation rates had previously been measured. Rotation rates from spot modulation are subject to measuring multiples of the true rotation rate, and also sometimes over-correct for this effect. This is not the case for asteroseismic rotation, and so some stars may appear at the 2:1 and 1:2 lines in the Figure. We fit a line of the form $P_{\rm surf} = m \times P_{\rm seis}$, using the larger of the asymmetrical uncertainties on asteroseismic rotation, and excluding stars above the 1.8:1 line to avoid outliers (transparent in the Figure). Our fit found a value of $m = 0.96 \pm 0.03$, showing a close agreement ($<2\sigma$) between asteroseismic and surface rotation rates on a population level. 
%Of the stars used in this comparison, 21 had a median value of $P_{\rm s}$ in the gyrochronology analysis. We repeated the model fit for these stars only, and found a value of $m = 0.96 \pm 0.04$.  
This overall agreement between spot and seismic rotation rates is in line with previous studies of Sun-like stars \cite{gizon+2013, chaplin+2013, nielsen+2015, benomar+2015}, and so we can conclude that our asteroseismic ensemble can be reliably used to draw inferences about gyrochronology.\\

\begin{figure}
	\centering
	\includegraphics[width=0.6\textwidth]{Images/surf-seis-comparison.pdf}
	\caption{Comparisons between asteroseismic and photometric measures of stellar rotation, for 48 stars. Literature values were taken, in order of priority, from: \cite{garcia+2014} and \cite{nielsen+2013} (i.e. if a value was reported in \cite{nielsen+2013} and also in \cite{garcia+2014}, the latter was used). Stars used in \cite{vansaders+2016} are highlighted. The four triangles represent stars included in the original \cite{nielsen+2015} work. The solid line indicates 1:1 line, while the dash-dotted lines represent 2:1 and 1:2 lines. The shaded region around the 1:1 line is the $1\sigma$ credible interval of the fit to the data, the result of which is shown as the title of the figure. Stars that were not included in the fitting process are transparent.}
	\label{fig:protlit}
\end{figure}

\textbf{Comparing gyrochronology models:} To evaluate the implications of our seismic ensemble for gyrochronology, we compared our new asteroseismic rotation rates to two population models of rotational evolution: a `standard' model, which assumes a traditional angular momentum transport through magnetically driven stellar winds \cite{skumanich1972, kawaler1988}, and one where weakened magnetic braking takes place (hereafter the `WMB model'). The WMB model was identical to the standard model in its input physics, except for the condition that angular momentum loss ceases above a critical Rossby number of $Ro_{\rm crit} = 1.97$. Instead of comparing our rotation rates directly to models of individual stars, we instead compared them to synthetic populations based on what the \textit{Kepler} field would look like under these two different braking laws \cite{girardi+2012,vansaders+2019}. For the full details on how these models were constructed, see \textbf{\textit{Methods}}.

In order to ensure our results were based on the most robust rotation measurements, we focused on the stars with the best convergence metrics in the mode frequency fitting\footnote{Excluding those with a number of effective samples $n_{\rm eff} < 25\%$ and a Gelman-Rubin metric of $\hat{R}>1.1$ \cite{gelman+rubin1992,salvatier+2016}} (see \textbf{\textit{Extended Data Table 1}}). The remaining sample of 73 stars contained 4 potential sub-giants, 22 `hot' stars and 47 main sequence stars. Figure \ref{fig:fullsample} shows these stars plotted over the top of both the standard and WMB population models in rotation-temperature space, alongside the stars not used in the gyrochronology analysis (shown without uncertainties).

\begin{figure*}
	\centering
	\includegraphics[width=\textwidth]{fullsample.png}
	\caption{Stars for which rotation was measured in this work, plotted over stellar population models evolved under standard (top) and weakened magnetic braking (WMB, bottom) prescriptions of angular momentum evolution. The 21 Stars that were excluded from the gyrochronology analysis are marked with crosses and without uncertainties (see text). Circles, squares and triangles denote main sequence, sub-giant, and `hot' stars respectively. Stars with a metallicity of $\bigl\lvert \feh \bigl\rvert > 0.4$ are coloured differently (darker), to indicate that they fall outside the functional range of the \cite{vansaders+2019} stellar models. Note that only period-temperature space is shown here, but when evaluating between the two model prescriptions we also considered mass, age and metallicity.}
	\label{fig:fullsample}
\end{figure*}

We evaluated our ensemble against both population models simultaneously using a Bayesian mixture model, which treated the data as being drawn from a mixture made from Kernel Density Estimates (KDEs) of both populations, modulated by a mixing parameter $P_{\rm s}$. In the limit that $P_{\rm s} \rightarrow 1$, the data are mostly likely drawn from the standard model. In the limit $P_{\rm s} \rightarrow 0$, the WMB model is more likely. This mixture model was evaluated in a five-dimensional parameter space, accounting for the population distributions in effective temperature ($T_{\rm eff}$), metallicity (\feh), asteroseismic mass ($M$) and age ($t$) and our new rotation periods ($P$).

We fit our mixture model to each star individually, which yielded a posterior probability density for the value of $P_{\rm s}$ for that star. In order to assess the posterior probability for the full sample, we multiplied the individual posterior probabilities for $P_{\rm s}$. This was done for the full ensemble of 73 stars, as well as for the three different stellar types. These joint probabilities can be seen in Figure \ref{fig:gyroresults} in the left-hand plot. The cumulative probability can be seen in the right-hand plot. In both of these plots, the left-hand side holds 98.4\% of the probability\footnote{When performing this analysis on all 91 stars for which seismic rotation was measured, this value is 96.6\%.}, indicating that the weakened magnetic braking model is preferred. When only considering the 47 main sequence stars, for which we expect the effects of weakened magnetic braking to be strongest, this rises to 99.2\%. \\

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{modelresults.pdf}
	\caption{Posterior estimates of the mixture model parameter $P_{\rm s}$, broken down by stellar classification. A value of $P_{\rm s}$ close to zero indicates that the data are more consistent with a rotational evolution that includes weakened magnetic braking, whereas a value close to 1 indicates that the data are more consistent with a standard rotational evolution scenario. \textit{Left:} the product of histograms of the posterior estimates for 73 stars, or subsets thereof for different stellar classifications. \textit{Right:} the cumulative posterior probability for 73 stars, or subsets thereof for different stellar types.}
	\label{fig:gyroresults}
\end{figure}

\textbf{Verifying consequences for gyrochronology:} In order to verify that our results presented a meaningful confirmation of weakened magnetic braking, we ensured that the same results were obtained when using different subsets of the ensemble, and when exploring potential biases. We found that the joint posterior distribution did not change significantly when excluding stars with a metallicity outside the model range (purple in Figure \ref{fig:fullsample}), or those with extremely small uncertainties on asteroseismic mass.Stars in our sample with known binary companions that may disrupt their rotational evolution were not found to contribute strongly to the joint posterior probability of $P_{\rm s}$.

The same was true when repeating the analysis with increased uncertainties on asteroseismic mass and age, or when shifting asteroseismic ages younger by their systematic uncertainties \cite{silvaaguirre+2015, silvaaguirre+2017}.  For a full discussion of these tests and other biases, see \textbf{\textit{Methods}}.\\

Section 5.2.5.

We leave the reader with the following conclusions:
\begin{enumerate}
	\item We found that our ensemble was $98.4\%$ more likely to be drawn from a weakened magnetic braking model than a standard model of rotational evolution \cite{vansaders+pinsonneault2013}, at a critical Rossby number of $Ro_{\rm{crit}} = 1.97$. This work expands upon previous analysis (\cite{vansaders+2019}) by including quiescent stars older than the Sun. This conclusion was found to be robust against choice of ensemble members and potentially underestimated asteroseismic systematic uncertainties. A comparison to other braking laws \cite{matt+2015} and other gyrochronology relations using our asteroseismic ensemble will take place in future work.

	\item We compared our new asteroseismic rotation rates with surface rotation measures from spot modulation. Our findings replicate those in similar studies \cite{nielsen+2015,benomar+2015}, in the sense that we find no statistically significant difference between seismic and spot modulation measures of stellar rotation in our ensemble.

	\item We present new techniques for asteroseismic `peak-bagging' frequency analyses, specifically: expressing well known empirical asteroseismic relations in a hierarchical modeling structure to account for small-scale effects not included in our model (such as acoustic glitches), and replacing uninformative empirical relations with Gaussian Process priors.

	\item The new asteroseismic rotation catalogue presented in this work will act as an entry point for more detailed studies, including comparisons between different braking laws and individually modeled values of stellar rotation at different critical Rossby numbers.
\end{enumerate}

Finally, in the near future our asteroseismic rotation catalogue will be further complemented; with high-precision surface rotation measurements and atmospheric parameters from spectroscopic surveys such as LAMOST \cite{deng+2012}, 4MOST \cite{dejong+2014} and WEAVE \cite{dalton+2014}; and new asteroseismic measurements of age from the \ktwo and \tess missions. With these surveys, large scale ensemble asteroseismology will continue to increase the possibilities for understanding gyrochronology.\\

All analysis performed as part of this work is stored in an online public repository\footnote{\url{https://github.com/ojhall94/malatium}}.\\

\textbf{Software}: \texttt{astropy} \cite{astropycollaboration+2013, astropycollaboration+2018}, \texttt{astroquery} \cite{ginsburg+2019}, \texttt{Lightkurve} \cite{lightkurvecollaboration+2018}, \texttt{corner} \cite{foreman-mackey2016}, \texttt{emcee} \cite{foreman-mackey+2013}, \href{https://docs.daft-pgm.org/}{\texttt{daft}}, \texttt{Stan} \cite{carpenter+2017}, \texttt{PyStan} \cite{vanhoey+2013}, \texttt{iPython} \cite{perez+granger2007}, \texttt{Jupyter Notebooks} \cite{kluyver+2016}, \texttt{numpy}  \cite{vanderwalt+2011}, \texttt{pandas} and \texttt{scipy} \cite{mckinney2010}, \texttt{matplotlib} \cite{hunter2007}, \texttt{seaborn} \cite{waskom+2018}, \texttt{statsmodels} \cite{seabold+perktold2010}, \texttt{PyMC3} \cite{salvatier+2016}, \texttt{theano} \cite{thetheanodevelopmentteam+2016}.\\

\textbf{Acknowledgments}:The authors would like to thank Sean Matt, Ellis Avallone, Alex Dixon, Warrick Ball and Brett Morris for helpful discussions. 
OJH, GRD and WJC acknowledge the support of the UK Science and Technology Facilities Council (STFC). 
JvS acknowledges support from the TESS Guest Investigator Program (80NSSC18K18584).
MBN acknowledges support from the UK Space Agency
This work has received funding from the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (CartographY GA. 804752)
Funding for the Stellar Astrophysics Centre is provided by The Danish National Research Foundation (Grant agreement no.: DNRF106). 
LA, AB and VS acknowledge funding from the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation program (grant agreement No. 682393 AWESoMeStars). AB also acknowledges the support of the College of Engineering, Mathematics and Physical Sciences at the University of Exeter.
RAG acknowledges the support from the PLATO and GOLF CNES grants.
JT acknowledges that support for this work was provided by NASA through the NASA Hubble Fellowship grant $\#51424$ awarded by the Space Telescope Science Institute, which is operated by the Association of Universities for Research in Astronomy, Inc., for NASA, under contract NAS5-26555.
The computations described in this paper were performed using the University of Birmingham's BlueBEAR HPC service.
This paper includes data collected by the Kepler mission and obtained from the MAST data archive at the Space Telescope Science Institute (STScI). Funding for the Kepler mission is provided by the NASA Science Mission Directorate. STScI is operated by the Association of Universities for Research in Astronomy, Inc., under NASA contract NAS 5–26555.\\

% The best way to enter references is to use BibTeX:
%\bibliographystyle{naturemag}
%\bibliography{library.bib} %

\textbf{Author Contributions:} O.J.H. led the project, with help from G.R.D., J.V.S., M.B.N and W.J.C.. J.V.S. also led the development of the stellar population models. M.N.L., R.A.G. and S.K. both provided data or stellar models and along with J.T. assessed the validity of our asteroseismic results. L.A., A.B. and V.S. provided the assessment of the theoretical implications of the gyrochronology results. All authors have contributed to the interpretation of the data and the results, and all discussed and provided comments for all drafts of the paper.

\textbf{Competing Interests:} The authors declare that they have no competing financial interests.

\textbf{Correspondence:} Correspondence and requests for materials should be addressed to O.J.H.. (email: oliver.hall@esa.int).

\clearpage

\input{methods.tex}

\input{si.tex}

\bibliographystyle{naturemag}
\bibliography{library.bib} %

\end{document}