### 2020-10-07 paper/natastron/bblmethods.tex

\begin{thebibliography}{10}
\expandafter\ifx\csname url\endcsname\relax
  \def\url#1{\texttt{#1}}\fi
\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
\providecommand{\bibinfo}[2]{#2}
\providecommand{\eprint}[2][]{\url{#2}}

\bibitem{m_handberg+lund2014}
\bibinfo{author}{Handberg, R.} \& \bibinfo{author}{Lund, M.~N.}
\newblock \bibinfo{title}{Automated preparation of {{Kepler}} time series of
  planet hosts for asteroseismic analysis}.
\newblock \emph{\bibinfo{journal}{Monthly Notices of the Royal Astronomical
  Society}} \textbf{\bibinfo{volume}{445}}, \bibinfo{pages}{2698--2709}
  (\bibinfo{year}{2014}).

\bibitem{m_garcia+2011}
\bibinfo{author}{Garc{\'i}a, R.~A.} \emph{et~al.}
\newblock \bibinfo{title}{Preparation of {{Kepler}} light curves for
  asteroseismic analyses}.
\newblock \emph{\bibinfo{journal}{Monthly Notices of the Royal Astronomical
  Society}} \textbf{\bibinfo{volume}{414}}, \bibinfo{pages}{L6--L10}
  (\bibinfo{year}{2011}).

\bibitem{m_silvaaguirre+2015}
\bibinfo{author}{Silva~Aguirre, V.} \emph{et~al.}
\newblock \bibinfo{title}{Ages and fundamental properties of {{Kepler}}
  exoplanet host stars from asteroseismology}.
\newblock \emph{\bibinfo{journal}{Monthly Notices of the Royal Astronomical
  Society}} \textbf{\bibinfo{volume}{452}}, \bibinfo{pages}{2127}
  (\bibinfo{year}{2015}).

\bibitem{m_silvaaguirre+2017}
\bibinfo{author}{Silva~Aguirre, V.} \emph{et~al.}
\newblock \bibinfo{title}{Standing on the {{Shoulders}} of {{Dwarfs}}: The
  {{Kepler Asteroseismic LEGACY Sample}}. {{II}}.{{Radii}}, {{Masses}}, and
  {{Ages}}}.
\newblock \emph{\bibinfo{journal}{The Astrophysical Journal}}
  \textbf{\bibinfo{volume}{835}}, \bibinfo{pages}{173} (\bibinfo{year}{2017}).

\bibitem{m_christensen-dalsgaard2008}
\bibinfo{author}{{Christensen-Dalsgaard}, J.}
\newblock \bibinfo{title}{{{ASTEC}}\textemdash the {{Aarhus STellar Evolution
  Code}}}.
\newblock \emph{\bibinfo{journal}{Astrophysics and Space Science}}
  \textbf{\bibinfo{volume}{316}}, \bibinfo{pages}{13} (\bibinfo{year}{2008}).

\bibitem{m_huber+2013a}
\bibinfo{author}{Huber, D.} \emph{et~al.}
\newblock \bibinfo{title}{Stellar {{Spin}}-{{Orbit Misalignment}} in a
  {{Multiplanet System}}}.
\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{342}},
  \bibinfo{pages}{331} (\bibinfo{year}{2013}).

\bibitem{m_buchhave+latham2015}
\bibinfo{author}{Buchhave, L.~A.} \& \bibinfo{author}{Latham, D.~W.}
\newblock \bibinfo{title}{The {{Metallicities}} of {{Stars}} with and without
  {{Transiting Planets}}}.
\newblock \emph{\bibinfo{journal}{The Astrophysical Journal}}
  \textbf{\bibinfo{volume}{808}}, \bibinfo{pages}{187} (\bibinfo{year}{2015}).

\bibitem{m_astropycollaboration+2013}
\bibinfo{author}{{Astropy Collaboration}} \emph{et~al.}
\newblock \bibinfo{title}{Astropy: {{A}} community {{Python}} package for
  astronomy}.
\newblock \emph{\bibinfo{journal}{Astronomy and Astrophysics}}
  \textbf{\bibinfo{volume}{558}}, \bibinfo{pages}{A33} (\bibinfo{year}{2013}).

\bibitem{m_astropycollaboration+2018}
\bibinfo{author}{{Astropy Collaboration}} \emph{et~al.}
\newblock \bibinfo{title}{The {{Astropy Project}}: {{Building}} an
  {{Open}}-science {{Project}} and {{Status}} of the v2.0 {{Core Package}}}.
\newblock \emph{\bibinfo{journal}{The Astronomical Journal}}
  \textbf{\bibinfo{volume}{156}}, \bibinfo{pages}{123} (\bibinfo{year}{2018}).

\bibitem{m_ginsburg+2019}
\bibinfo{author}{Ginsburg, A.} \emph{et~al.}
\newblock \bibinfo{title}{Astroquery: {{An Astronomical Web}}-querying
  {{Package}} in {{Python}}}.
\newblock \emph{\bibinfo{journal}{The Astronomical Journal}}
  \textbf{\bibinfo{volume}{157}}, \bibinfo{pages}{98} (\bibinfo{year}{2019}).

\bibitem{m_mckinney2010}
\bibinfo{author}{McKinney, W.}
\newblock \bibinfo{title}{Data {{Structures}} for {{Statistical Computing}} in
  {{Python}}}.
\newblock In \emph{\bibinfo{booktitle}{Proceedings of the 9th {{Python}} in
  {{Science Conference}}}}, \bibinfo{pages}{51--56} (\bibinfo{year}{2010}).

\bibitem{m_davies+2015}
\bibinfo{author}{Davies, G.~R.} \emph{et~al.}
\newblock \bibinfo{title}{Asteroseismic inference on rotation, gyrochronology
  and planetary system dynamics of 16 {{Cygni}}}.
\newblock \emph{\bibinfo{journal}{Monthly Notices of the Royal Astronomical
  Society}} \textbf{\bibinfo{volume}{446}}, \bibinfo{pages}{2959}
  (\bibinfo{year}{2015}).

\bibitem{m_harvey1985}
\bibinfo{author}{Harvey, J.}
\newblock \bibinfo{title}{High-{{Resolution Helioseismology}}}.
\newblock \emph{\bibinfo{journal}{Future Missions in Solar, Heliospheric \&
  Space Plasma Physics}} \textbf{\bibinfo{volume}{235}}, \bibinfo{pages}{199}
  (\bibinfo{year}{1985}).

\bibitem{m_tassoul1980}
\bibinfo{author}{Tassoul, M.}
\newblock \bibinfo{title}{Asymptotic approximations for stellar nonradial
  pulsations}.
\newblock \emph{\bibinfo{journal}{The Astrophysical Journal Supplement Series}}
  \textbf{\bibinfo{volume}{43}}, \bibinfo{pages}{469} (\bibinfo{year}{1980}).

\bibitem{m_vrard+2016}
\bibinfo{author}{Vrard, M.}, \bibinfo{author}{Mosser, B.} \&
  \bibinfo{author}{Samadi, R.}
\newblock \bibinfo{title}{Period spacings in red giants. {{II}}. {{Automated}}
  measurement}.
\newblock \emph{\bibinfo{journal}{Astronomy and Astrophysics}}
  \textbf{\bibinfo{volume}{588}}, \bibinfo{pages}{A87} (\bibinfo{year}{2016}).

\bibitem{m_hogg+2010}
\bibinfo{author}{Hogg, D.~W.}, \bibinfo{author}{Bovy, J.} \&
  \bibinfo{author}{Lang, D.}
\newblock \bibinfo{title}{Data analysis recipes: {{Fitting}} a model to data}.
\newblock \emph{\bibinfo{journal}{ArXiv e-prints}}
  \bibinfo{pages}{arXiv:1008.4686} (\bibinfo{year}{2010}).

\bibitem{m_hall+2019}
\bibinfo{author}{Hall, O.~J.} \emph{et~al.}
\newblock \bibinfo{title}{Testing asteroseismology with {{Gaia DR2}}:
  {{Hierarchical}} models of the {{Red Clump}}}.
\newblock \emph{\bibinfo{journal}{Monthly Notices of the Royal Astronomical
  Society}} \textbf{\bibinfo{volume}{486}}, \bibinfo{pages}{3569--3585}
  (\bibinfo{year}{2019}).
\newblock \eprint{1904.07919}.

\bibitem{m_mazumdar+2014}
\bibinfo{author}{Mazumdar, A.} \emph{et~al.}
\newblock \bibinfo{title}{Measurement of acoustic glitches in solar-type stars
  from oscillation frequencies observed by {{Kepler}}}.
\newblock \emph{\bibinfo{journal}{The Astrophysical Journal}}
  \textbf{\bibinfo{volume}{782}}, \bibinfo{pages}{18} (\bibinfo{year}{2014}).
\newblock \eprint{1312.4907}.

\bibitem{m_chaplin+basu2017}
\bibinfo{author}{Chaplin, W.~J.} \& \bibinfo{author}{Basu}.
\newblock \emph{\bibinfo{title}{Asteroseismic {{Data Analysis}}:
  {{Foundations}} and {{Techniques}}}} (\bibinfo{publisher}{{Princeton
  University Press}}, \bibinfo{address}{{Princeton, New Jersey}},
  \bibinfo{year}{2017}), \bibinfo{edition}{1st} edn.

\bibitem{m_vanhoey+2013}
\bibinfo{author}{Van~Hoey, S.}, \bibinfo{author}{{van der Kwast}, J.},
  \bibinfo{author}{Nopens, I.} \& \bibinfo{author}{Seuntjens, P.}
\newblock \bibinfo{title}{Python package for model {{STructure ANalysis}}
  ({{pySTAN}})}.
\newblock \emph{\bibinfo{journal}{EGU General Assembly Conference Abstracts}}
  \textbf{\bibinfo{volume}{15}}, \bibinfo{pages}{EGU2013--10059}
  (\bibinfo{year}{2013}).

\bibitem{m_carpenter+2017}
\bibinfo{author}{Carpenter, B.} \emph{et~al.}
\newblock \bibinfo{title}{Stan: {{A Probabilistic Programming Language}}}.
\newblock \emph{\bibinfo{journal}{Journal of Statistical Software}}
  \textbf{\bibinfo{volume}{76}}, \bibinfo{pages}{1--32} (\bibinfo{year}{2017}).

\bibitem{m_vanderwalt+2011}
\bibinfo{author}{{van der Walt}, S.}, \bibinfo{author}{Colbert, S.~C.} \&
  \bibinfo{author}{Varoquaux, G.}
\newblock \bibinfo{title}{The {{NumPy Array}}: {{A Structure}} for {{Efficient
  Numerical Computation}}}.
\newblock \emph{\bibinfo{journal}{Computing in Science and Engineering}}
  \textbf{\bibinfo{volume}{13}}, \bibinfo{pages}{22--30}
  (\bibinfo{year}{2011}).

\bibitem{m_salvatier+2016}
\bibinfo{author}{Salvatier, J.}, \bibinfo{author}{Wiecki, T.~V.} \&
  \bibinfo{author}{Fonnesbeck, C.}
\newblock \bibinfo{title}{Probabilistic programming in {{Python}} using
  {{PyMC3}}}.
\newblock \emph{\bibinfo{journal}{PeerJ Computer Science}}
  \textbf{\bibinfo{volume}{2}}, \bibinfo{pages}{e55} (\bibinfo{year}{2016}).

\bibitem{m_thetheanodevelopmentteam+2016}
\bibinfo{author}{{The Theano Development Team}} \emph{et~al.}
\newblock \bibinfo{title}{Theano: {{A Python}} framework for fast computation
  of mathematical expressions}.
\newblock \emph{\bibinfo{journal}{arXiv:1605.02688 [cs]}}
  (\bibinfo{year}{2016}).
\newblock \eprint{1605.02688}.

\bibitem{m_gelman+rubin1992}
\bibinfo{author}{Gelman, A.} \& \bibinfo{author}{Rubin, D.~B.}
\newblock \bibinfo{title}{Inference from {{Iterative Simulation Using Multiple
  Sequences}}}.
\newblock \emph{\bibinfo{journal}{Statistical Science}}
  \textbf{\bibinfo{volume}{7}}, \bibinfo{pages}{457--472}
  (\bibinfo{year}{1992}).

\bibitem{m_vansaders+pinsonneault2013}
\bibinfo{author}{{van Saders}, J.~L.} \& \bibinfo{author}{Pinsonneault, M.~H.}
\newblock \bibinfo{title}{Fast {{Star}}, {{Slow Star}}; {{Old Star}}, {{Young
  Star}}: {{Subgiant Rotation}} as a {{Population}} and {{Stellar Physics
  Diagnostic}}}.
\newblock \emph{\bibinfo{journal}{The Astrophysical Journal}}
  \textbf{\bibinfo{volume}{776}}, \bibinfo{pages}{67} (\bibinfo{year}{2013}).

\bibitem{m_vansaders+2016}
\bibinfo{author}{{van Saders}, J.~L.} \emph{et~al.}
\newblock \bibinfo{title}{Weakened magnetic braking as the origin of
  anomalously rapid rotation in old field stars}.
\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{529}},
  \bibinfo{pages}{181} (\bibinfo{year}{2016}).

\bibitem{m_vansaders+2019}
\bibinfo{author}{{van Saders}, J.~L.}, \bibinfo{author}{Pinsonneault, M.~H.} \&
  \bibinfo{author}{Barbieri, M.}
\newblock \bibinfo{title}{Forward {{Modeling}} of the {{Kepler Stellar Rotation
  Period Distribution}}: {{Interpreting Periods}} from {{Mixed}} and {{Biased
  Stellar Populations}}}.
\newblock \emph{\bibinfo{journal}{The Astrophysical Journal}}
  \textbf{\bibinfo{volume}{872}}, \bibinfo{pages}{128} (\bibinfo{year}{2019}).

\bibitem{m_girardi+2012}
\bibinfo{author}{Girardi, L.} \emph{et~al.}
\newblock \bibinfo{title}{{{TRILEGAL}}, a {{TRIdimensional modeL}} of {{thE
  GALaxy}}: {{Status}} and {{Future}}}.
\newblock \emph{\bibinfo{journal}{Astrophysics and Space Science Proceedings}}
  \textbf{\bibinfo{volume}{26}}, \bibinfo{pages}{165} (\bibinfo{year}{2012}).

\bibitem{m_berger+2020}
\bibinfo{author}{Berger, T.~A.} \emph{et~al.}
\newblock \bibinfo{title}{The \${{Gaia}}\$-\${{Kepler}}\$ {{Stellar Properties
  Catalog I}}: {{Homogeneous Fundamental Properties}} for 186,000
  \${{Kepler}}\$ {{Stars}}}.
\newblock \emph{\bibinfo{journal}{arXiv:2001.07737 [astro-ph]}}
  (\bibinfo{year}{2020}).
\newblock \eprint{2001.07737}.

\bibitem{m_gaiacollaboration+2018}
\bibinfo{author}{{Gaia Collaboration}} \emph{et~al.}
\newblock \bibinfo{title}{Gaia {{Data Release}} 2. {{Summary}} of the contents
  and survey properties}.
\newblock \emph{\bibinfo{journal}{ArXiv e-prints}}
  \bibinfo{pages}{arXiv:1804.09365} (\bibinfo{year}{2018}).

\bibitem{m_raghavan+2010}
\bibinfo{author}{Raghavan, D.} \emph{et~al.}
\newblock \bibinfo{title}{A {{Survey}} of {{Stellar Families}}:
  {{Multiplicity}} of {{Solar}}-type {{Stars}}}.
\newblock \emph{\bibinfo{journal}{The Astrophysical Journal Supplement Series}}
  \textbf{\bibinfo{volume}{190}}, \bibinfo{pages}{1--42}
  (\bibinfo{year}{2010}).

\bibitem{m_chaplin+2011}
\bibinfo{author}{Chaplin, W.~J.} \emph{et~al.}
\newblock \bibinfo{title}{Ensemble {{Asteroseismology}} of {{Solar}}-{{Type
  Stars}} with the {{NASA Kepler Mission}}}.
\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{332}},
  \bibinfo{pages}{213} (\bibinfo{year}{2011}).

\bibitem{m_seabold+perktold2010}
\bibinfo{author}{Seabold, S.} \& \bibinfo{author}{Perktold, J.}
\newblock \bibinfo{title}{Statsmodels: {{Econometric}} and {{Statistical
  Modeling}} with {{Python}}} \bibinfo{pages}{5} (\bibinfo{year}{2010}).

\bibitem{m_betancourt+girolami2013}
\bibinfo{author}{Betancourt, M.~J.} \& \bibinfo{author}{Girolami, M.}
\newblock \bibinfo{title}{Hamiltonian {{Monte Carlo}} for {{Hierarchical
  Models}}}.
\newblock \emph{\bibinfo{journal}{arXiv:1312.0906 [stat]}}
  (\bibinfo{year}{2013}).
\newblock \eprint{1312.0906}.

\bibitem{m_foreman-mackey+2013}
\bibinfo{author}{{Foreman-Mackey}, D.}, \bibinfo{author}{Hogg, D.~W.},
  \bibinfo{author}{Lang, D.} \& \bibinfo{author}{Goodman, J.}
\newblock \bibinfo{title}{Emcee: {{The MCMC Hammer}}}.
\newblock \emph{\bibinfo{journal}{Publications of the Astronomical Society of
  the Pacific}} \textbf{\bibinfo{volume}{125}}, \bibinfo{pages}{306--312}
  (\bibinfo{year}{2013}).
\newblock \eprint{1202.3665}.

\bibitem{m_foreman-mackey2016}
\bibinfo{author}{{Foreman-Mackey}, D.}
\newblock \bibinfo{title}{Corner.py: {{Scatterplot}} matrices in {{Python}}}.
\newblock \emph{\bibinfo{journal}{The Journal of Open Source Software}}
  \textbf{\bibinfo{volume}{1}} (\bibinfo{year}{2016}).

\bibitem{m_nielsen+2015}
\bibinfo{author}{Nielsen, M.~B.}, \bibinfo{author}{Schunker, H.},
  \bibinfo{author}{Gizon, L.} \& \bibinfo{author}{Ball, W.~H.}
\newblock \bibinfo{title}{Constraining differential rotation of {{Sun}}-like
  stars from asteroseismic and starspot rotation periods}.
\newblock \emph{\bibinfo{journal}{Astronomy and Astrophysics}}
  \textbf{\bibinfo{volume}{582}}, \bibinfo{pages}{A10} (\bibinfo{year}{2015}).

\bibitem{m_benomar+2018}
\bibinfo{author}{Benomar, O.} \emph{et~al.}
\newblock \bibinfo{title}{Asteroseismic detection of latitudinal differential
  rotation in 13 {{Sun}}-like stars}.
\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{361}},
  \bibinfo{pages}{1231} (\bibinfo{year}{2018}).

\bibitem{m_beck2000}
\bibinfo{author}{Beck, J.~G.}
\newblock \bibinfo{title}{A comparison of differential rotation measurements -
  ({{Invited Review}})}.
\newblock \emph{\bibinfo{journal}{Solar Physics}}
  \textbf{\bibinfo{volume}{191}}, \bibinfo{pages}{47--70}
  (\bibinfo{year}{2000}).

\bibitem{m_benomar+2015}
\bibinfo{author}{Benomar, O.}, \bibinfo{author}{Takata, M.},
  \bibinfo{author}{Shibahashi, H.}, \bibinfo{author}{Ceillier, T.} \&
  \bibinfo{author}{Garc{\'i}a, R.~A.}
\newblock \bibinfo{title}{Nearly uniform internal rotation of solar-like
  main-sequence stars revealed by space-based asteroseismology and
  spectroscopic measurements}.
\newblock \emph{\bibinfo{journal}{Monthly Notices of the Royal Astronomical
  Society}} \textbf{\bibinfo{volume}{452}}, \bibinfo{pages}{2654--2674}
  (\bibinfo{year}{2015}).

\bibitem{m_doyle+2014}
\bibinfo{author}{Doyle, A.~P.}, \bibinfo{author}{Davies, G.~R.},
  \bibinfo{author}{Smalley, B.}, \bibinfo{author}{Chaplin, W.~J.} \&
  \bibinfo{author}{Elsworth, Y.}
\newblock \bibinfo{title}{Determining stellar macroturbulence using
  asteroseismic rotational velocities from {{Kepler}}}.
\newblock \emph{\bibinfo{journal}{Monthly Notices of the Royal Astronomical
  Society}} \textbf{\bibinfo{volume}{444}}, \bibinfo{pages}{3592--3602}
  (\bibinfo{year}{2014}).

\bibitem{m_tayar+2015}
\bibinfo{author}{Tayar, J.} \emph{et~al.}
\newblock \bibinfo{title}{Rapid {{Rotation}} of {{Low}}-mass {{Red Giants Using
  APOKASC}}: {{A Measure}} of {{Interaction Rates}} on the
  {{Post}}-main-sequence}.
\newblock \emph{\bibinfo{journal}{The Astrophysical Journal}}
  \textbf{\bibinfo{volume}{807}}, \bibinfo{pages}{82} (\bibinfo{year}{2015}).

\bibitem{m_schofield+2019}
\bibinfo{author}{Schofield, M.} \emph{et~al.}
\newblock \bibinfo{title}{The {{Asteroseismic Target List}} for {{Solar}}-like
  {{Oscillators Observed}} in 2 minute {{Cadence}} with the {{Transiting
  Exoplanet Survey Satellite}}}.
\newblock \emph{\bibinfo{journal}{The Astrophysical Journal Supplement Series}}
  \textbf{\bibinfo{volume}{241}}, \bibinfo{pages}{12} (\bibinfo{year}{2019}).

\bibitem{m_mathur+2019}
\bibinfo{author}{{Mathur}, S.} \emph{et~al.}
\newblock \bibinfo{title}{{Revisiting the impact of stellar magnetic activity
  on the detection of solar-like oscillations by Kepler}}.
\newblock \emph{\bibinfo{journal}{Frontiers in Astronomy and Space Sciences}}
  \textbf{\bibinfo{volume}{6}}, \bibinfo{pages}{46} (\bibinfo{year}{2019}).
\newblock \eprint{1907.01415}.

\bibitem{m_amard+matt2020}
\bibinfo{author}{Amard, L.} \& \bibinfo{author}{Matt, S.~P.}
\newblock \bibinfo{title}{The {{Impact}} of {{Metallicity}} on the
  {{Evolution}} of {{Rotation}} and {{Magnetic Activity}} of {{Sun}}-{{Like
  Stars}}}.
\newblock \emph{\bibinfo{journal}{arXiv:2001.10404 [astro-ph]}}
  (\bibinfo{year}{2020}).
\newblock \eprint{2001.10404}.

\bibitem{m_matt+2015}
\bibinfo{author}{Matt, S.~P.}, \bibinfo{author}{Brun, A.~S.},
  \bibinfo{author}{Baraffe, I.}, \bibinfo{author}{Bouvier, J.} \&
  \bibinfo{author}{Chabrier, G.}
\newblock \bibinfo{title}{The {{Mass}}-dependence of {{Angular Momentum
  Evolution}} in {{Sun}}-like {{Stars}}}.
\newblock \emph{\bibinfo{journal}{The Astrophysical Journal}}
  \textbf{\bibinfo{volume}{799}}, \bibinfo{pages}{L23} (\bibinfo{year}{2015}).

\bibitem{m_leiner+2019}
\bibinfo{author}{Leiner, E.}, \bibinfo{author}{Mathieu, R.~D.},
  \bibinfo{author}{Vanderburg, A.}, \bibinfo{author}{Gosnell, N.~M.} \&
  \bibinfo{author}{Smith, J.~C.}
\newblock \bibinfo{title}{Blue {{Lurkers}}: {{Hidden Blue Stragglers}} on the
  {{M67 Main Sequence Identified}} from {{Their Kepler}}/{{K2 Rotation
  Periods}}}.
\newblock \emph{\bibinfo{journal}{The Astrophysical Journal}}
  \textbf{\bibinfo{volume}{881}}, \bibinfo{pages}{47} (\bibinfo{year}{2019}).

\bibitem{m_fleming+2019}
\bibinfo{author}{Fleming, D.~P.}, \bibinfo{author}{Barnes, R.},
  \bibinfo{author}{Davenport, J. R.~A.} \& \bibinfo{author}{Luger, R.}
\newblock \bibinfo{title}{Rotation {{Period Evolution}} in {{Low}}-mass
  {{Binary Stars}}: {{The Impact}} of {{Tidal Torques}} and {{Magnetic
  Braking}}}.
\newblock \emph{\bibinfo{journal}{The Astrophysical Journal}}
  \textbf{\bibinfo{volume}{881}}, \bibinfo{pages}{88} (\bibinfo{year}{2019}).

\bibitem{m_maxted+2015}
\bibinfo{author}{Maxted, P. F.~L.}, \bibinfo{author}{Serenelli, A.~M.} \&
  \bibinfo{author}{Southworth, J.}
\newblock \bibinfo{title}{Comparison of gyrochronological and isochronal age
  estimates for transiting exoplanet host stars}.
\newblock \emph{\bibinfo{journal}{Astronomy and Astrophysics}}
  \textbf{\bibinfo{volume}{577}}, \bibinfo{pages}{A90} (\bibinfo{year}{2015}).

\bibitem{m_gallet+delorme2019}
\bibinfo{author}{Gallet, F.} \& \bibinfo{author}{Delorme, P.}
\newblock \bibinfo{title}{Star-planet tidal interaction and the limits of
  gyrochronology}.
\newblock \emph{\bibinfo{journal}{arXiv e-prints}}
  \textbf{\bibinfo{volume}{1905}}, \bibinfo{pages}{arXiv:1905.06070}
  (\bibinfo{year}{2019}).

\bibitem{m_benbakoura+2019}
\bibinfo{author}{Benbakoura, M.}, \bibinfo{author}{R{\'e}ville, V.},
  \bibinfo{author}{Brun, A.~S.}, \bibinfo{author}{{Le Poncin-Lafitte}, C.} \&
  \bibinfo{author}{Mathis, S.}
\newblock \bibinfo{title}{Evolution of star-planet systems under magnetic
  braking and tidal interaction}.
\newblock \emph{\bibinfo{journal}{Astronomy and Astrophysics}}
  \textbf{\bibinfo{volume}{621}}, \bibinfo{pages}{A124} (\bibinfo{year}{2019}).

\bibitem{m_ceillier+2016}
\bibinfo{author}{Ceillier, T.} \emph{et~al.}
\newblock \bibinfo{title}{Rotation periods and seismic ages of {{KOIs}} -
  comparison with stars without detected planets from {{Kepler}} observations}.
\newblock \emph{\bibinfo{journal}{Monthly Notices of the Royal Astronomical
  Society}} \textbf{\bibinfo{volume}{456}}, \bibinfo{pages}{119--125}
  (\bibinfo{year}{2016}).

\end{thebibliography}