### 2020-10-09 paper/natastron/response/coverletter.tex

\documentclass[11pt, a4paper, sans, colorlinks, linkcolor=True]{moderncv}
\moderncvstyle{classic}
\moderncvcolor{black}
\usepackage{ragged2e}
\usepackage[utf8]{inputenc}
\usepackage{enumitem}
\usepackage[left=1in, right=1in, bottom=0.5in, top=0.5in]{geometry}
%\addtolength{\topmargin}{-.875in}
\pagenumbering{gobble}
\usepackage{setspace}

%Personal Details
\name{Oliver James}{Hall}
\title{Coverletter}
\address{European Space Agency (ESA)\\ European Space Research and Technology Centre (ESTEC)\\ Keplerlaan 1}{2201 AZ Noordwijk}{The Netherlands}
\email{ojhall94@gmail.com}
\homepage{https://www.ojhall94.github.io}

\begin{document}
\definecolor{links}{HTML}{3f28bd}
\hypersetup{urlcolor=links}
\recipient{\ }{\ }
\date{\today}
\opening{Dear Dr. Woods,}
\makelettertitle
\justify
\vspace{-0.5cm}

We hope this letter finds you well. We have completed our response to the reviewer, which is attached below this cover letter (and uploaded separately with resubmission). Based on our discussion over email on the reformatting of the original draft, we have done our best to reshuffle the paper's contents. We think the updated draft best matches the Nature format while keeping it accessible to the reader. However the Methods is longer than conventional, so please let us know if you have any suggestions on what we should move to the Supplementary Information, if necessary.

We have also added data and code availability statements. At the moment, our main results table is in Extended Data, as we don't think it is necessary to include it in the main text. Ideally we would publish this table digitally along with the article (as well as host it on Vizier). If you have any input on whether this table should be (partially) included in the main text, we would be grateful also.

The word count breakdown is as follows:
\begin{itemize}
	\item Abstract: 196
	\item Introduction: 500
	\item Main body (incl. Introduction): 2721
	\item Methods: 5145
	\item Supplementary Information: 2544
	\item Extended Data: Figures and Table only.
\end{itemize}

\vspace{0.5cm}

Yours sincerely, \\
\vspace{0.5em}
\includegraphics[scale=0.2]{signature.png} \\ 
\vspace{0.5em}
\textbf{Oliver J. Hall}, on behalf of the authors. \\
\end{document}