### 2021-02-15 paper/natastron/Publication/bblmethods.tex
tidying up for publication

\begin{thebibliography}{10}
\expandafter\ifx\csname url\endcsname\relax
\def\url#1{\texttt{#1}}\fi
\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
\providecommand{\bibinfo}[2]{#2}
\providecommand{\eprint}[2][]{\url{#2}}

\bibitem{handberg+lund2014}
\bibinfo{author}{Handberg, R.} \& \bibinfo{author}{Lund, M.~N.}
\newblock \bibinfo{title}{Automated preparation of {{Kepler}} time series of
  planet hosts for asteroseismic analysis}.
\newblock \emph{\bibinfo{journal}{Monthly Notices of the Royal Astronomical
  Society}} \textbf{\bibinfo{volume}{445}}, \bibinfo{pages}{2698--2709}
  (\bibinfo{year}{2014}).

\bibitem{garcia+2011}
\bibinfo{author}{Garc{\'i}a, R.~A.} \emph{et~al.}
\newblock \bibinfo{title}{Preparation of {{Kepler}} light curves for
  asteroseismic analyses}.
\newblock \emph{\bibinfo{journal}{Monthly Notices of the Royal Astronomical
  Society}} \textbf{\bibinfo{volume}{414}}, \bibinfo{pages}{L6--L10}
  (\bibinfo{year}{2011}).

\bibitem{huber+2013a}
\bibinfo{author}{Huber, D.} \emph{et~al.}
\newblock \bibinfo{title}{Stellar {{Spin}}-{{Orbit Misalignment}} in a
  {{Multiplanet System}}}.
\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{342}},
  \bibinfo{pages}{331} (\bibinfo{year}{2013}).

\bibitem{buchhave+latham2015}
\bibinfo{author}{Buchhave, L.~A.} \& \bibinfo{author}{Latham, D.~W.}
\newblock \bibinfo{title}{The {{Metallicities}} of {{Stars}} with and without
  {{Transiting Planets}}}.
\newblock \emph{\bibinfo{journal}{The Astrophysical Journal}}
  \textbf{\bibinfo{volume}{808}}, \bibinfo{pages}{187} (\bibinfo{year}{2015}).

\bibitem{astropycollaboration+2013}
\bibinfo{author}{{Astropy Collaboration}} \emph{et~al.}
\newblock \bibinfo{title}{Astropy: {{A}} community {{Python}} package for
  astronomy}.
\newblock \emph{\bibinfo{journal}{Astronomy and Astrophysics}}
  \textbf{\bibinfo{volume}{558}}, \bibinfo{pages}{A33} (\bibinfo{year}{2013}).

\bibitem{mckinney2010}
\bibinfo{author}{McKinney, W.}
\newblock \bibinfo{title}{Data {{Structures}} for {{Statistical Computing}} in
  {{Python}}}.
\newblock In \emph{\bibinfo{booktitle}{Proceedings of the 9th {{Python}} in
  {{Science Conference}}}}, \bibinfo{pages}{51--56} (\bibinfo{year}{2010}).

\bibitem{harvey1985}
\bibinfo{author}{Harvey, J.}
\newblock \bibinfo{title}{High-{{Resolution Helioseismology}}}.
\newblock \emph{\bibinfo{journal}{Future Missions in Solar, Heliospheric \&
  Space Plasma Physics}} \textbf{\bibinfo{volume}{235}}, \bibinfo{pages}{199}
  (\bibinfo{year}{1985}).

\bibitem{tassoul1980}
\bibinfo{author}{Tassoul, M.}
\newblock \bibinfo{title}{Asymptotic approximations for stellar nonradial
  pulsations}.
\newblock \emph{\bibinfo{journal}{The Astrophysical Journal Supplement Series}}
  \textbf{\bibinfo{volume}{43}}, \bibinfo{pages}{469} (\bibinfo{year}{1980}).

\bibitem{hogg+2010}
\bibinfo{author}{Hogg, D.~W.}, \bibinfo{author}{Bovy, J.} \&
  \bibinfo{author}{Lang, D.}
\newblock \bibinfo{title}{Data analysis recipes: {{Fitting}} a model to data}.
\newblock \emph{\bibinfo{journal}{ArXiv e-prints}}
  \bibinfo{pages}{arXiv:1008.4686} (\bibinfo{year}{2010}).

\bibitem{mazumdar+2014}
\bibinfo{author}{Mazumdar, A.} \emph{et~al.}
\newblock \bibinfo{title}{Measurement of acoustic glitches in solar-type stars
  from oscillation frequencies observed by {{Kepler}}}.
\newblock \emph{\bibinfo{journal}{The Astrophysical Journal}}
  \textbf{\bibinfo{volume}{782}}, \bibinfo{pages}{18} (\bibinfo{year}{2014}).
\newblock \eprint{1312.4907}.

\bibitem{chaplin+basu2017}
\bibinfo{author}{Chaplin, W.~J.} \& \bibinfo{author}{Basu}.
\newblock \emph{\bibinfo{title}{Asteroseismic {{Data Analysis}}:
  {{Foundations}} and {{Techniques}}}} (\bibinfo{publisher}{{Princeton
  University Press}}, \bibinfo{address}{{Princeton, New Jersey}},
  \bibinfo{year}{2017}), \bibinfo{edition}{1st} edn.

\bibitem{vanhoey+2013}
\bibinfo{author}{Van~Hoey, S.}, \bibinfo{author}{{van der Kwast}, J.},
  \bibinfo{author}{Nopens, I.} \& \bibinfo{author}{Seuntjens, P.}
\newblock \bibinfo{title}{Python package for model {{STructure ANalysis}}
  ({{pySTAN}})}.
\newblock \emph{\bibinfo{journal}{EGU General Assembly Conference Abstracts}}
  \textbf{\bibinfo{volume}{15}}, \bibinfo{pages}{EGU2013--10059}
  (\bibinfo{year}{2013}).

\bibitem{vanderwalt+2011}
\bibinfo{author}{{van der Walt}, S.}, \bibinfo{author}{Colbert, S.~C.} \&
  \bibinfo{author}{Varoquaux, G.}
\newblock \bibinfo{title}{The {{NumPy Array}}: {{A Structure}} for {{Efficient
  Numerical Computation}}}.
\newblock \emph{\bibinfo{journal}{Computing in Science and Engineering}}
  \textbf{\bibinfo{volume}{13}}, \bibinfo{pages}{22--30}
  (\bibinfo{year}{2011}).

\bibitem{gaiacollaboration+2018}
\bibinfo{author}{{Gaia Collaboration}} \emph{et~al.}
\newblock \bibinfo{title}{Gaia {{Data Release}} 2. {{Summary}} of the contents
  and survey properties}.
\newblock \emph{\bibinfo{journal}{ArXiv e-prints}}
  \bibinfo{pages}{arXiv:1804.09365} (\bibinfo{year}{2018}).

\bibitem{raghavan+2010}
\bibinfo{author}{Raghavan, D.} \emph{et~al.}
\newblock \bibinfo{title}{A {{Survey}} of {{Stellar Families}}:
  {{Multiplicity}} of {{Solar}}-type {{Stars}}}.
\newblock \emph{\bibinfo{journal}{The Astrophysical Journal Supplement Series}}
  \textbf{\bibinfo{volume}{190}}, \bibinfo{pages}{1--42}
  (\bibinfo{year}{2010}).

\bibitem{chaplin+2011}
\bibinfo{author}{Chaplin, W.~J.} \emph{et~al.}
\newblock \bibinfo{title}{Ensemble {{Asteroseismology}} of {{Solar}}-{{Type
  Stars}} with the {{NASA Kepler Mission}}}.
\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{332}},
  \bibinfo{pages}{213} (\bibinfo{year}{2011}).

\bibitem{seabold+perktold2010}
\bibinfo{author}{Seabold, S.} \& \bibinfo{author}{Perktold, J.}
\newblock \bibinfo{title}{Statsmodels: {{Econometric}} and {{Statistical
  Modeling}} with {{Python}}} \bibinfo{pages}{5} (\bibinfo{year}{2010}).

\bibitem{foreman-mackey+2013}
\bibinfo{author}{{Foreman-Mackey}, D.}, \bibinfo{author}{Hogg, D.~W.},
  \bibinfo{author}{Lang, D.} \& \bibinfo{author}{Goodman, J.}
\newblock \bibinfo{title}{Emcee: {{The MCMC Hammer}}}.
\newblock \emph{\bibinfo{journal}{Publications of the Astronomical Society of
  the Pacific}} \textbf{\bibinfo{volume}{125}}, \bibinfo{pages}{306--312}
  (\bibinfo{year}{2013}).
\newblock \eprint{1202.3665}.

\end{thebibliography}