### 2021-02-15 paper/natastron/Publication/bblpaper.tex

\begin{thebibliography}{10}
	\expandafter\ifx\csname url\endcsname\relax
	\def\url#1{\texttt{#1}}\fi
	\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
	\providecommand{\bibinfo}[2]{#2}
	\providecommand{\eprint}[2][]{\url{#2}}

	\bibitem{barnes2007}
	\bibinfo{author}{Barnes, S.~A.}
	\newblock \bibinfo{title}{Ages for {{Illustrative Field Stars Using
				Gyrochronology}}: {{Viability}}, {{Limitations}}, and {{Errors}}}.
	\newblock \emph{\bibinfo{journal}{Astrophys. J.}}
	\textbf{\bibinfo{volume}{669}}, \bibinfo{pages}{1167} (\bibinfo{year}{2007}).

	\bibitem{meibom+2015}
	\bibinfo{author}{Meibom, S.} \emph{et~al.}
	\newblock \bibinfo{title}{A spin-down clock for cool stars from observations of
		a 2.5-billion-year-old cluster}.
	\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{517}},
	\bibinfo{pages}{589} (\bibinfo{year}{2015}).

	\bibitem{leiner+2019}
	\bibinfo{author}{Leiner, E.}, \bibinfo{author}{Mathieu, R.~D.},
	\bibinfo{author}{Vanderburg, A.}, \bibinfo{author}{Gosnell, N.~M.} \&
	\bibinfo{author}{Smith, J.~C.}
	\newblock \bibinfo{title}{Blue {{Lurkers}}: {{Hidden Blue Stragglers}} on the
		{{M67 Main Sequence Identified}} from {{Their Kepler}}/{{K2 Rotation
				Periods}}}.
	\newblock \emph{\bibinfo{journal}{Astrophys. J.}}
	\textbf{\bibinfo{volume}{881}}, \bibinfo{pages}{47} (\bibinfo{year}{2019}).

	\bibitem{claytor+2019}
	\bibinfo{author}{{Claytor}, Z.~R.} \emph{et~al.}
	\newblock \bibinfo{title}{{Chemical Evolution in the Milky Way: Rotation-based
			Ages for APOGEE-Kepler Cool Dwarf Stars}}.
	\newblock \emph{\bibinfo{journal}{Astrophys. J.}}
	\textbf{\bibinfo{volume}{888}}, \bibinfo{pages}{43} (\bibinfo{year}{2020}).
	\newblock \eprint{1911.04518}.

	\bibitem{barnes+2016}
	\bibinfo{author}{Barnes, S.~A.}, \bibinfo{author}{Weingrill, J.},
	\bibinfo{author}{Fritzewski, D.}, \bibinfo{author}{Strassmeier, K.~G.} \&
	\bibinfo{author}{Platais, I.}
	\newblock \bibinfo{title}{Rotation {{Periods}} for {{Cool Stars}} in the 4
		{{Gyr}} old {{Open Cluster M67}}, {{The Solar}}-{{Stellar Connection}}, and
		the {{Applicability}} of {{Gyrochronology}} to at least {{Solar Age}}}.
	\newblock \emph{\bibinfo{journal}{Astrophys. J.}}
	\textbf{\bibinfo{volume}{823}}, \bibinfo{pages}{16} (\bibinfo{year}{2016}).

	\bibitem{borucki+2010}
	\bibinfo{author}{Borucki, W.~J.}, \bibinfo{author}{Koch, D.} \&
	\bibinfo{author}{Team, K.~S.}
	\newblock \bibinfo{title}{Kepler {{Planet Detection Mission}}: {{Highlights}}
		of the {{First Results}}}.
	\newblock \emph{\bibinfo{journal}{AAS/Division for Planetary Sciences Meeting
			Abstracts \#42}} \textbf{\bibinfo{volume}{42}}, \bibinfo{pages}{47.03}
	(\bibinfo{year}{2010}).

	\bibitem{silvaaguirre+2015}
	\bibinfo{author}{Silva~Aguirre, V.} \emph{et~al.}
	\newblock \bibinfo{title}{Ages and fundamental properties of {{Kepler}}
		exoplanet host stars from asteroseismology}.
	\newblock \emph{\bibinfo{journal}{Mon. Not. R. Astron. Soc.}}
	\textbf{\bibinfo{volume}{452}}, \bibinfo{pages}{2127} (\bibinfo{year}{2015}).

	\bibitem{angus+2015}
	\bibinfo{author}{Angus, R.}, \bibinfo{author}{Aigrain, S.},
	\bibinfo{author}{{Foreman-Mackey}, D.} \& \bibinfo{author}{McQuillan, A.}
	\newblock \bibinfo{title}{Calibrating gyrochronology using {{Kepler}}
		asteroseismic targets}.
	\newblock \emph{\bibinfo{journal}{Mon. Not. R. Astron. Soc.}}
	\textbf{\bibinfo{volume}{450}}, \bibinfo{pages}{1787} (\bibinfo{year}{2015}).

	\bibitem{nielsen+2015}
	\bibinfo{author}{Nielsen, M.~B.}, \bibinfo{author}{Schunker, H.},
	\bibinfo{author}{Gizon, L.} \& \bibinfo{author}{Ball, W.~H.}
	\newblock \bibinfo{title}{Constraining differential rotation of {{Sun}}-like
		stars from asteroseismic and starspot rotation periods}.
	\newblock \emph{\bibinfo{journal}{Astron. Astrophys.}}
	\textbf{\bibinfo{volume}{582}}, \bibinfo{pages}{A10} (\bibinfo{year}{2015}).

	\bibitem{davies+2015}
	\bibinfo{author}{Davies, G.~R.} \emph{et~al.}
	\newblock \bibinfo{title}{Asteroseismic inference on rotation, gyrochronology
		and planetary system dynamics of 16 {{Cygni}}}.
	\newblock \emph{\bibinfo{journal}{Mon. Not. R. Astron. Soc.}}
	\textbf{\bibinfo{volume}{446}}, \bibinfo{pages}{2959} (\bibinfo{year}{2015}).

	\bibitem{vansaders+2016}
	\bibinfo{author}{{van Saders}, J.~L.} \emph{et~al.}
	\newblock \bibinfo{title}{Weakened magnetic braking as the origin of
		anomalously rapid rotation in old field stars}.
	\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{529}},
	\bibinfo{pages}{181} (\bibinfo{year}{2016}).

	\bibitem{reville+2015}
	\bibinfo{author}{{R{\'e}ville}, V.}, \bibinfo{author}{{Brun}, A.~S.},
	\bibinfo{author}{{Matt}, S.~P.}, \bibinfo{author}{{Strugarek}, A.} \&
	\bibinfo{author}{{Pinto}, R.~F.}
	\newblock \bibinfo{title}{{The Effect of Magnetic Topology on Thermally Driven
			Wind: Toward a General Formulation of the Braking Law}}.
	\newblock \emph{\bibinfo{journal}{Astrophys. J.}}
	\textbf{\bibinfo{volume}{798}}, \bibinfo{pages}{116} (\bibinfo{year}{2015}).
	\newblock \eprint{1410.8746}.

	\bibitem{garraffo+2016}
	\bibinfo{author}{Garraffo, C.}, \bibinfo{author}{Drake, J.~J.} \&
	\bibinfo{author}{Cohen, O.}
	\newblock \bibinfo{title}{The {{Missing Magnetic Morphology Term}} in {{Stellar
				Rotation Evolution}}}.
	\newblock \emph{\bibinfo{journal}{Astronomy \& Astrophysics}}
	\textbf{\bibinfo{volume}{595}}, \bibinfo{pages}{A110} (\bibinfo{year}{2016}).
	\newblock \eprint{1607.06096}.

	\bibitem{metcalfe+2016}
	\bibinfo{author}{Metcalfe, T.~S.}, \bibinfo{author}{Egeland, R.} \&
	\bibinfo{author}{{van Saders}, J.}
	\newblock \bibinfo{title}{Stellar {{Evidence That}} the {{Solar Dynamo May Be}}
		in {{Transition}}}.
	\newblock \emph{\bibinfo{journal}{Astrophys. J. Letters}}
	\textbf{\bibinfo{volume}{826}}, \bibinfo{pages}{L2} (\bibinfo{year}{2016}).

	\bibitem{metcalfe+2019}
	\bibinfo{author}{{Metcalfe}, T.~S.} \emph{et~al.}
	\newblock \bibinfo{title}{{LBT/PEPSI Spectropolarimetry of a Magnetic
			Morphology Shift in Old Solar-type Stars}}.
	\newblock \emph{\bibinfo{journal}{Astrophys. J.}}
	\textbf{\bibinfo{volume}{887}}, \bibinfo{pages}{L38} (\bibinfo{year}{2019}).
	\newblock \eprint{1912.01186}.

	\bibitem{see+2019}
	\bibinfo{author}{See, V.} \emph{et~al.}
	\newblock \bibinfo{title}{Do {{Non}}-dipolar {{Magnetic Fields Contribute}} to
		{{Spin}}-down {{Torques}}?}
	\newblock \emph{\bibinfo{journal}{Astrophys. J.}}
	\textbf{\bibinfo{volume}{886}}, \bibinfo{pages}{120} (\bibinfo{year}{2019}).

	\bibitem{mcquillan+2014}
	\bibinfo{author}{McQuillan, A.}, \bibinfo{author}{Mazeh, T.} \&
	\bibinfo{author}{Aigrain, S.}
	\newblock \bibinfo{title}{Rotation {{Periods}} of 34,030 {{Kepler
				Main}}-{{Sequence Stars}}: {{The Full Autocorrelation Sample}}}.
	\newblock \emph{\bibinfo{journal}{Astrophys. J. Supplement Series}}
	\textbf{\bibinfo{volume}{211}}, \bibinfo{pages}{24} (\bibinfo{year}{2014}).
	\newblock \eprint{1402.5694}.

	\bibitem{matt+2015}
	\bibinfo{author}{Matt, S.~P.}, \bibinfo{author}{Brun, A.~S.},
	\bibinfo{author}{Baraffe, I.}, \bibinfo{author}{Bouvier, J.} \&
	\bibinfo{author}{Chabrier, G.}
	\newblock \bibinfo{title}{The {{Mass}}-dependence of {{Angular Momentum
				Evolution}} in {{Sun}}-like {{Stars}}}.
	\newblock \emph{\bibinfo{journal}{Astrophys. J.}}
	\textbf{\bibinfo{volume}{799}}, \bibinfo{pages}{L23} (\bibinfo{year}{2015}).

	\bibitem{reinhold+2020}
	\bibinfo{author}{{Reinhold}, T.} \emph{et~al.}
	\newblock \bibinfo{title}{{The Sun is less active than other solar-like
			stars}}.
	\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{368}},
	\bibinfo{pages}{518--521} (\bibinfo{year}{2020}).
	\newblock \eprint{2005.01401}.

	\bibitem{vansaders+2019}
	\bibinfo{author}{{van Saders}, J.~L.}, \bibinfo{author}{Pinsonneault, M.~H.} \&
	\bibinfo{author}{Barbieri, M.}
	\newblock \bibinfo{title}{Forward {{Modeling}} of the {{Kepler Stellar Rotation
				Period Distribution}}: {{Interpreting Periods}} from {{Mixed}} and {{Biased
				Stellar Populations}}}.
	\newblock \emph{\bibinfo{journal}{Astrophys. J.}}
	\textbf{\bibinfo{volume}{872}}, \bibinfo{pages}{128} (\bibinfo{year}{2019}).

	\bibitem{ledoux1951}
	\bibinfo{author}{Ledoux, P.}
	\newblock \bibinfo{title}{The {{Nonradial Oscillations}} of {{Gaseous Stars}}
		and the {{Problem}} of {{Beta Canis Majoris}}.}
	\newblock \emph{\bibinfo{journal}{Astrophys. J.}}
	\textbf{\bibinfo{volume}{114}}, \bibinfo{pages}{373} (\bibinfo{year}{1951}).

	\bibitem{davies+2016}
	\bibinfo{author}{Davies, G.~R.} \emph{et~al.}
	\newblock \bibinfo{title}{Oscillation frequencies for 35 {{Kepler}} solar-type
		planet-hosting stars using {{Bayesian}} techniques and machine learning}.
	\newblock \emph{\bibinfo{journal}{Mon. Not. R. Astron. Soc.}}
	\textbf{\bibinfo{volume}{456}}, \bibinfo{pages}{2183} (\bibinfo{year}{2016}).

	\bibitem{lund+2017}
	\bibinfo{author}{Lund, M.~N.} \emph{et~al.}
	\newblock \bibinfo{title}{Standing on the {{Shoulders}} of {{Dwarfs}}: The
		{{Kepler Asteroseismic LEGACY Sample}}. {{I}}. {{Oscillation Mode
				Parameters}}}.
	\newblock \emph{\bibinfo{journal}{Astrophys. J.}}
	\textbf{\bibinfo{volume}{835}}, \bibinfo{pages}{172} (\bibinfo{year}{2017}).

	\bibitem{silvaaguirre+2017}
	\bibinfo{author}{Silva~Aguirre, V.} \emph{et~al.}
	\newblock \bibinfo{title}{Standing on the {{Shoulders}} of {{Dwarfs}}: The
		{{Kepler Asteroseismic LEGACY Sample}}. {{II}}.{{Radii}}, {{Masses}}, and
		{{Ages}}}.
	\newblock \emph{\bibinfo{journal}{Astrophys. J.}}
	\textbf{\bibinfo{volume}{835}}, \bibinfo{pages}{173} (\bibinfo{year}{2017}).

	\bibitem{garcia+2014}
	\bibinfo{author}{Garc{\'i}a, R.~A.} \emph{et~al.}
	\newblock \bibinfo{title}{Rotation and magnetism of {{Kepler}} pulsating
		solar-like stars. {{Towards}} asteroseismically calibrated age-rotation
		relations}.
	\newblock \emph{\bibinfo{journal}{Astron. Astrophys.}}
	\textbf{\bibinfo{volume}{572}}, \bibinfo{pages}{A34} (\bibinfo{year}{2014}).

	\bibitem{kraft1967}
	\bibinfo{author}{Kraft, R.~P.}
	\newblock \bibinfo{title}{Studies of {{Stellar Rotation}}. {{V}}. {{The
				Dependence}} of {{Rotation}} on {{Age}} among {{Solar}}-{{Type Stars}}}.
	\newblock \emph{\bibinfo{journal}{Astrophys. J.}}
	\textbf{\bibinfo{volume}{150}}, \bibinfo{pages}{551} (\bibinfo{year}{1967}).

	\bibitem{bedding+2010}
	\bibinfo{author}{Bedding, T.~R.} \emph{et~al.}
	\newblock \bibinfo{title}{Solar-like {{Oscillations}} in {{Low}}-luminosity
		{{Red Giants}}: {{First Results}} from {{Kepler}}}.
	\newblock \emph{\bibinfo{journal}{Astrophys. J. Letters}}
	\textbf{\bibinfo{volume}{713}}, \bibinfo{pages}{L176--L181}
	(\bibinfo{year}{2010}).

	\bibitem{bonanno+frohlich2015}
	\bibinfo{author}{Bonanno, A.} \& \bibinfo{author}{Fr{\"o}hlich, H.-E.}
	\newblock \bibinfo{title}{A {{Bayesian}} estimation of the helioseismic solar
		age}.
	\newblock \emph{\bibinfo{journal}{Astron. Astrophys.}}
	\textbf{\bibinfo{volume}{580}}, \bibinfo{pages}{A130} (\bibinfo{year}{2015}).

	\bibitem{paxton+2017}
	\bibinfo{author}{{Paxton}, B.} \emph{et~al.}
	\newblock \bibinfo{title}{{Modules for Experiments in Stellar Astrophysics
			(MESA): Convective Boundaries, Element Diffusion, and Massive Star
			Explosions}}.
	\newblock \emph{\bibinfo{journal}{Astrophys. J. Supplement Series}}
	\textbf{\bibinfo{volume}{234}}, \bibinfo{pages}{34} (\bibinfo{year}{2018}).
	\newblock \eprint{1710.08424}.

	\bibitem{hunter2007}
	\bibinfo{author}{Hunter, J.~D.}
	\newblock \bibinfo{title}{Matplotlib: {{A 2D Graphics Environment}}}.
	\newblock \emph{\bibinfo{journal}{Computing in Science and Engineering}}
	\textbf{\bibinfo{volume}{9}}, \bibinfo{pages}{90--95} (\bibinfo{year}{2007}).

	\bibitem{benomar+2018}
	\bibinfo{author}{Benomar, O.} \emph{et~al.}
	\newblock \bibinfo{title}{Asteroseismic detection of latitudinal differential
		rotation in 13 {{Sun}}-like stars}.
	\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{361}},
	\bibinfo{pages}{1231} (\bibinfo{year}{2018}).

	\bibitem{lund+2014}
	\bibinfo{author}{Lund, M.~N.}, \bibinfo{author}{Miesch, M.~S.} \&
	\bibinfo{author}{{Christensen-Dalsgaard}, J.}
	\newblock \bibinfo{title}{Differential rotation in main-sequence solar-like
		stars: {{Qualitative}} inference from asteroseismic data}.
	\newblock \emph{\bibinfo{journal}{Astrophys. J.}}
	\textbf{\bibinfo{volume}{790}}, \bibinfo{pages}{121} (\bibinfo{year}{2014}).
	\newblock \eprint{1406.7873}.

	\bibitem{benomar+2015}
	\bibinfo{author}{Benomar, O.}, \bibinfo{author}{Takata, M.},
	\bibinfo{author}{Shibahashi, H.}, \bibinfo{author}{Ceillier, T.} \&
	\bibinfo{author}{Garc{\'i}a, R.~A.}
	\newblock \bibinfo{title}{Nearly uniform internal rotation of solar-like
		main-sequence stars revealed by space-based asteroseismology and
		spectroscopic measurements}.
	\newblock \emph{\bibinfo{journal}{Mon. Not. R. Astron. Soc.}}
	\textbf{\bibinfo{volume}{452}}, \bibinfo{pages}{2654--2674}
	(\bibinfo{year}{2015}).

	\bibitem{gizon+2013}
	\bibinfo{author}{Gizon, L.} \emph{et~al.}
	\newblock \bibinfo{title}{Seismic constraints on rotation of {{Sun}}-like star
		and mass of exoplanet}.
	\newblock \emph{\bibinfo{journal}{Proceedings of the National Academy of
			Science}} \textbf{\bibinfo{volume}{110}}, \bibinfo{pages}{13267}
	(\bibinfo{year}{2013}).

	\bibitem{chaplin+2013}
	\bibinfo{author}{Chaplin, W.~J.} \emph{et~al.}
	\newblock \bibinfo{title}{Asteroseismic {{Determination}} of {{Obliquities}} of
		the {{Exoplanet Systems Kepler}}-50 and {{Kepler}}-65}.
	\newblock \emph{\bibinfo{journal}{Astrophys. J.}}
	\textbf{\bibinfo{volume}{766}}, \bibinfo{pages}{101} (\bibinfo{year}{2013}).

	\bibitem{nielsen+2013}
	\bibinfo{author}{Nielsen, M.~B.}, \bibinfo{author}{Gizon, L.},
	\bibinfo{author}{Schunker, H.} \& \bibinfo{author}{Karoff, C.}
	\newblock \bibinfo{title}{Rotation periods of 12 000 main-sequence {{Kepler}}
		stars: {{Dependence}} on stellar spectral type and comparison with v sin i
		observations}.
	\newblock \emph{\bibinfo{journal}{Astron. Astrophys.}}
	\textbf{\bibinfo{volume}{557}}, \bibinfo{pages}{L10} (\bibinfo{year}{2013}).

	\bibitem{skumanich1972}
	\bibinfo{author}{{Skumanich}, A.}
	\newblock \bibinfo{title}{{Time Scales for Ca II Emission Decay, Rotational
			Braking, and Lithium Depletion}}.
	\newblock \emph{\bibinfo{journal}{Astrophys. J.}}
	\textbf{\bibinfo{volume}{171}}, \bibinfo{pages}{565} (\bibinfo{year}{1972}).

	\bibitem{kawaler1988}
	\bibinfo{author}{Kawaler, S.~D.}
	\newblock \bibinfo{title}{Angular momentum loss in low-mass stars}.
	\newblock \emph{\bibinfo{journal}{Astrophys. J.}}
	\textbf{\bibinfo{volume}{333}}, \bibinfo{pages}{236} (\bibinfo{year}{1988}).

	\bibitem{girardi+2012}
	\bibinfo{author}{Girardi, L.} \emph{et~al.}
	\newblock \bibinfo{title}{{{TRILEGAL}}, a {{TRIdimensional modeL}} of {{thE
				GALaxy}}: {{Status}} and {{Future}}}.
	\newblock \emph{\bibinfo{journal}{Astrophysics and Space Science Proceedings}}
	\textbf{\bibinfo{volume}{26}}, \bibinfo{pages}{165} (\bibinfo{year}{2012}).

	\bibitem{berger+2020}
	\bibinfo{author}{{Berger}, T.~A.}, \bibinfo{author}{{Huber}, D.},
	\bibinfo{author}{{Gaidos}, E.}, \bibinfo{author}{{van Saders}, J.~L.} \&
	\bibinfo{author}{{Weiss}, L.~M.}
	\newblock \bibinfo{title}{{The Gaia-Kepler Stellar Properties Catalog. II.
			Planet Radius Demographics as a Function of Stellar Mass and Age}}.
	\newblock \emph{\bibinfo{journal}{Astron. J.}} \textbf{\bibinfo{volume}{160}},
	\bibinfo{pages}{108} (\bibinfo{year}{2020}).
	\newblock \eprint{2005.14671}.

	\bibitem{gelman+rubin1992}
	\bibinfo{author}{Gelman, A.} \& \bibinfo{author}{Rubin, D.~B.}
	\newblock \bibinfo{title}{Inference from {{Iterative Simulation Using Multiple
				Sequences}}}.
	\newblock \emph{\bibinfo{journal}{Statistical Science}}
	\textbf{\bibinfo{volume}{7}}, \bibinfo{pages}{457--472}
	(\bibinfo{year}{1992}).

	\bibitem{salvatier+2016}
	\bibinfo{author}{Salvatier, J.}, \bibinfo{author}{Wiecki, T.~V.} \&
	\bibinfo{author}{Fonnesbeck, C.}
	\newblock \bibinfo{title}{Probabilistic programming in {{Python}} using
		{{PyMC3}}}.
	\newblock \emph{\bibinfo{journal}{PeerJ Computer Science}}
	\textbf{\bibinfo{volume}{2}}, \bibinfo{pages}{e55} (\bibinfo{year}{2016}).

	\bibitem{metcalfe+egeland2019}
	\bibinfo{author}{Metcalfe, T.~S.} \& \bibinfo{author}{Egeland, R.}
	\newblock \bibinfo{title}{Understanding the {{Limitations}} of
		{{Gyrochronology}} for {{Old Field Stars}}}.
	\newblock \emph{\bibinfo{journal}{Astrophys. J.}}
	\textbf{\bibinfo{volume}{871}}, \bibinfo{pages}{39} (\bibinfo{year}{2019}).
	\newblock \eprint{1811.11905}.

	\bibitem{lorenzo-oliveira+2019}
	\bibinfo{author}{{Lorenzo-Oliveira}, D.} \emph{et~al.}
	\newblock \bibinfo{title}{Constraining the evolution of stellar rotation using
		solar twins}.
	\newblock \emph{\bibinfo{journal}{Mon. Not. R. Astron. Soc.}}
	\textbf{\bibinfo{volume}{485}}, \bibinfo{pages}{L68} (\bibinfo{year}{2019}).

	\bibitem{vansaders+pinsonneault2013}
	\bibinfo{author}{{van Saders}, J.~L.} \& \bibinfo{author}{Pinsonneault, M.~H.}
	\newblock \bibinfo{title}{Fast {{Star}}, {{Slow Star}}; {{Old Star}}, {{Young
				Star}}: {{Subgiant Rotation}} as a {{Population}} and {{Stellar Physics
				Diagnostic}}}.
	\newblock \emph{\bibinfo{journal}{Astrophys. J.}}
	\textbf{\bibinfo{volume}{776}}, \bibinfo{pages}{67} (\bibinfo{year}{2013}).

	\bibitem{barnes2010}
	\bibinfo{author}{Barnes, S.~A.}
	\newblock \bibinfo{title}{A {{Simple Nonlinear Model}} for the {{Rotation}} of
		{{Main}}-sequence {{Cool Stars}}. {{I}}. {{Introduction}}, {{Implications}}
		for {{Gyrochronology}}, and {{Color}}-{{Period Diagrams}}}.
	\newblock \emph{\bibinfo{journal}{Astrophys. J.}}
	\textbf{\bibinfo{volume}{722}}, \bibinfo{pages}{222} (\bibinfo{year}{2010}).

	\bibitem{deng+2012}
	\bibinfo{author}{Deng, L.-C.} \emph{et~al.}
	\newblock \bibinfo{title}{{{LAMOST Experiment}} for {{Galactic Understanding}}
		and {{Exploration}} ({{LEGUE}}) \textemdash{} {{The}} survey's science plan}.
	\newblock \emph{\bibinfo{journal}{Research in Astron. Astrophys.}}
	\textbf{\bibinfo{volume}{12}}, \bibinfo{pages}{735--754}
	(\bibinfo{year}{2012}).

	\bibitem{dejong+2014}
	\bibinfo{author}{{de Jong}, R.~S.} \emph{et~al.}
	\newblock \bibinfo{title}{{4MOST: 4-metre Multi-Object Spectroscopic
			Telescope}}.
	\newblock In \bibinfo{editor}{{Ramsay}, S.~K.}, \bibinfo{editor}{{McLean},
		I.~S.} \& \bibinfo{editor}{{Takami}, H.} (eds.)
	\emph{\bibinfo{booktitle}{Ground-based and Airborne Instrumentation for
			Astronomy V}}, vol. \bibinfo{volume}{9147} of \emph{\bibinfo{series}{Society
			of Photo-Optical Instrumentation Engineers (SPIE) Conference Series}},
	\bibinfo{pages}{91470M} (\bibinfo{year}{2014}).

	\bibitem{dalton+2014}
	\bibinfo{author}{{Dalton}, G.} \emph{et~al.}
	\newblock \bibinfo{title}{{Project overview and update on WEAVE: the next
			generation wide-field spectroscopy facility for the William Herschel
			Telescope}}.
	\newblock In \bibinfo{editor}{{Ramsay}, S.~K.}, \bibinfo{editor}{{McLean},
		I.~S.} \& \bibinfo{editor}{{Takami}, H.} (eds.)
	\emph{\bibinfo{booktitle}{Ground-based and Airborne Instrumentation for
			Astronomy V}}, vol. \bibinfo{volume}{9147} of \emph{\bibinfo{series}{Society
			of Photo-Optical Instrumentation Engineers (SPIE) Conference Series}},
	\bibinfo{pages}{91470L} (\bibinfo{year}{2014}).
	\newblock \eprint{1412.0843}.

	\bibitem{blanton+2019}
	\bibinfo{author}{{Blanton}, M.} \emph{et~al.}
	\newblock \bibinfo{title}{{The Sloan Digital Sky Survey as an Archetypal
			Mid-Scale Program}}.
	\newblock In \emph{\bibinfo{booktitle}{Bull. Am. Astron. Soc.}},
	vol.~\bibinfo{volume}{51}, \bibinfo{pages}{196} (\bibinfo{year}{2019}).

	\bibitem{kollmeier+2019}
	\bibinfo{author}{{Kollmeier}, J.} \emph{et~al.}
	\newblock \bibinfo{title}{{SDSS-V Pioneering Panoptic Spectroscopy}}.
	\newblock In \emph{\bibinfo{booktitle}{Bull. Am. Astron. Soc.}},
	vol.~\bibinfo{volume}{51}, \bibinfo{pages}{274} (\bibinfo{year}{2019}).

\end{thebibliography}