### 2021-02-16 paper/natastron/Publication/coverletter.tex
done coverletter

\documentclass[11pt, a4paper, sans, colorlinks, linkcolor=True]{moderncv}
\moderncvstyle{classic}
\moderncvcolor{black}
\usepackage{ragged2e}
\usepackage[utf8]{inputenc}
%\usepackage{enumitem}
\usepackage[left=1in, right=1in, bottom=0.5in, top=0.5in]{geometry}
%\addtolength{\topmargin}{-.875in}
\pagenumbering{gobble}
\usepackage{setspace}

%Personal Details
\name{Oliver James}{Hall}
\title{Coverletter}
\address{European Space Agency (ESA)\\ European Space Research and Technology Centre (ESTEC)\\ Keplerlaan 1}{2201 AZ Noordwijk}{The Netherlands}
\email{ojhall94@gmail.com}
\homepage{https://www.ojhall94.github.io}

\begin{document}
\definecolor{links}{HTML}{3f28bd}
\hypersetup{urlcolor=links}
\recipient{\ }{\ }
\date{\today}
\opening{Dear Dr. Woods,}
\makelettertitle
\justify
\vspace{-0.5cm}

We have made the requested changes, including those discussed over email, to the best of my ability. Please don't hesitate to get in touch if anything is missing!

For the purposes of advertising this work on Twitter, my handle is @asteronomer.

Those of other authors involved (that have Twitter) are:
\begin{itemize}
	\item Guy R. Davies: @CitationWarrior\\
	\item Mikkel N. Lund: @miklnl41\\
	\item Saniya Khan: @rilasani\\
	\item Victor See: @AstroVictorSee\\
	\item Jamie Tayar: @StellarTayar\\
\end{itemize}

I'll leave it up to you whether you want to them. The institutional handles are, in order of appearance on the manuscript:

\begin{itemize}
	\item @ESAscience\\
	\item @unibirmingham\\
	\item @sacscience\\
	\item @UHIfA\\
	\item @CEAIrfu\\
	\item @UniofExeter\\
\end{itemize}

Yours sincerely, \\
\vspace{0.5em}
\includegraphics[scale=0.2]{signature.png} \\ 
\vspace{0.5em}
\textbf{Oliver J. Hall}, on behalf of the authors. \\
\end{document}